---
title: "Coffee Shop Survey Report"
author: "Kabamba Mutale"
date: "2025-04-22"
output:
  word_document: 
    toc: true
  html_document: default
  pdf_document: 
    toc: true
    number_sections: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# **Install & Load Libraries**

Install packages by running **install.packages(c("dplyr", "ggplot2", "readr"))** in the console. This tells R to install the listed packages in one go first if they are not, thereafter, you just load them.

```{r}
# Load libraries and set global options
library(readxl)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(rmarkdown)
library(tinytex)
```

# **Import Data**

Use `read_excel` to load the dataset by providing the path for your dataset.

```{r}
#Loading Data
data <- read_excel("J:/r/coffee_shop_survey.xlsx")
```

# **Inspecting the Data**

**View the structure**

```{r}
# View the structure 
str(data)
```

**View the first few rows**

```{r}
# View the first few rows
head(data)
```

# **Cleaning Data**

**Check for missing values**

```{r}
# Check for missing values
colSums(is.na(data))
```

**Remove rows with any NA values (optional)**

```{r}
# Remove rows with any NA values (optional)
data2 <- na.omit(data)
```

**Changing column name (optional)**

```{r}
# changing column name
names(data2)[names(data2) == "Time_Spent (min)"] <- "Time_Spent_min"
```

**Confirming column name change (optional)**

```{r}
# Check column names
colnames(data2)
```

# **Descriptive Analysis**

### **1. What is the average age of customers visiting the coffee shop?**

```{r}
# Average Age of Customers

average_age <- mean(data2$Age,na.rm = TRUE)
print(paste("Average Age:", round(average_age, 1)))

```

**Visual KPI for Average Age**

```{r}
#Create a data frame for ploting
age_data <- data.frame(Metric = "Average Age", Value = average_age)

#Plot

ggplot(data.frame(Age = average_age), aes(x = 1, y = Age)) + geom_text(aes(label = paste0("Average Age\n", round(Age, 1))), size = 8, fontface = "bold") + theme_void() + theme( plot.background = element_rect(fill = "#f5f5f5", color = NA), panel.background = element_rect(fill = "#f5f5f5", color = NA), plot.title = element_text(hjust = 0.5, face = "bold", size = 14) ) + labs(title = "Key Metric")
```

### **2. What is the gender distribution among respondents?**

```{r}
gender_distribution <- data2 %>% 
  count(Gender, name = "Count")
  print(gender_distribution)
table(data2$Gender)
```

***Plot - Gender Distribution***

```{r}
# Gender Distribution among respondents
ggplot(gender_distribution, aes(x = Gender, y = Count, fill = Gender)) +
  geom_col() +
  labs(
      title = "Gender Distribution",
      x = "Gender of Customers",
      y = "Number Of Customers") + 
      theme_minimal() 
```

### **3. What are the most and least popular products?**

```{r}
# Most & Least Popular Products
product_counts <-
table(data2$Favorite_Product)
most_popular <-
names(product_counts[which.max(product_counts)])
least_popular <-
names(product_counts[which.min(product_counts)])
```

***Most Popular***

```{r}
print(paste("Most Popular Product:", most_popular))
```

***Least Popular***

```{r}
print(paste("Least Popular Product:", least_popular))
```

**Plot** - **Most & Least Popular Products**

```{r}
# Create a data frame from the table
product_counts <- as.data.frame(table(data$Favorite_Product))
colnames(product_counts) <- c("Product", "Count")
# Plot
ggplot(product_counts, aes(x = reorder(Product, -Count), y = Count, fill = Product)) +
  geom_bar(stat = "identity") +
  labs(title = "Favorite Product Popularity",
       x = "Product", y = "Number of Customers") +
  theme_minimal() +
  theme(legend.position = "none") +
  geom_text(aes(label = Count), vjust = -0.5)
```

# **Relationship & Comparison Questions**

### **1. Does visit frequency differ by gender or age group?**

***By Gender***

```{r}
#By Gender
aggregate(Visit_Frequency ~ Gender, data2, FUN=mean)
data2_summary <- data2 %>%
  group_by(Gender) %>%
  summarize(Mean_Visit_Frequency = mean(Visit_Frequency, na.rm = TRUE))

ggplot(data2_summary, aes(x = Gender, y = Mean_Visit_Frequency, fill = Gender)) +
  geom_col() +
  labs(
    title = "Mean Visit Frequency by Gender",
    x = "Gender",
    y = "Mean Visit Frequency"
  ) +
  theme_minimal()
         
```

***By Age***

```{r}
#By Age
data2$Age_Group <- cut(data2$Age, breaks=c(0, 17, 29, 44, 59, 100), labels=c("Child", "Young Adult", "Adult", "Old", "Senior"))
aggregate(Visit_Frequency ~ Age_Group, data2, FUN=mean)

ggplot(data2, aes(x = Age_Group, y = Visit_Frequency, fill = Age_Group)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Mean Visit Frequency by Age Group",
    x = "Age Group",
    y = "Mean Visit Frequency"
  ) +
  theme_minimal()

```

### **2. Is there a relationship between visit frequency and satisfaction score?**

***Correlation***

```{r}
cor(data2$Visit_Frequency, data2$Satisfaction_Score)
```

***Plot - Correlation***

```{r}
ggplot(data, aes(x = Visit_Frequency, y = Satisfaction_Score)) + geom_jitter(color = "purple") + geom_smooth(method = "lm", se = FALSE, color = "black") + labs(title = "Visit Frequency vs Satisfaction Score") + theme_minimal()
```

### **3. Do loyalty members spend more time at the coffee shop than non-members?**

```{r}
# First, summarize the data

summary_data <- data2%>% group_by(Loyalty_Member) %>% summarise(Avg_Time_Spent = mean(Time_Spent_min, na.rm = TRUE))

# Then plot

ggplot(summary_data, aes(x = factor(Loyalty_Member, levels = c("Yes", "No"), labels = c("Loyal", "Not Loyal")), y = Avg_Time_Spent, fill = Loyalty_Member)) + geom_col() + geom_text(aes(label = round(Avg_Time_Spent, 1)), vjust = -0.5, color = "black") + labs(title = "Time Spent by Loyalty Membership", y = "Time Spent (min)", x = "Loyalty Member") + theme_minimal()

t.test(Time_Spent_min ~ Loyalty_Member, data = data2)
```

The t-test indicates that loyalty program members tend to spend a bit more time in the coffee shop compared to non-members (average: 35.67 minutes vs. 30.53 minutes). However, the p-value (0.07917) exceeds the typical 0.05 significance level, implying that although there appears to be a slight difference, the evidence isn't strong enough to confidently conclude that the pattern is real.

```{r}
chi_square_data <- table(data2$Would_Recommend, data2$Loyalty_Member)
print(chi_square_data)

chi_square_df <- as.data.frame(chi_square_data)
colnames(chi_square_df) <- c("Would_Recommend", "Loyalty_Member", "Count")

chi_square_df$Loyalty_Member <- factor(chi_square_df$Loyalty_Member,
                                       levels = c("No", "Yes"),
                                       labels = c("Not Loyal", "Loyal"))

ggplot(chi_square_df, aes(x = Would_Recommend, y = Count, fill = Loyalty_Member)) +
  geom_col(position = "dodge") +  # "dodge" makes bars side-by-side instead of stacked
  geom_text(aes(label = Count), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, 
            size = 3.5) +  # Add labels on top of bars
  labs(title = "Customer Recommendation vs Loyalty Membership",
       x = "Would Recommend the Coffee Shop",
       y = "Number of Customers",
       fill = "Loyalty Member") +
  theme_minimal()


```

The chart shows the relationship between customer recommendation and loyalty membership status. Among those who would not recommend the coffee shop, loyal members outnumber non-loyal members (20 vs. 7). Among those who would recommend it, the numbers are more balanced (35 loyal vs. 38 not loyal). The Chi-Square test (p = 0.035) indicates a statistically significant association between loyalty membership and likelihood to recommend the coffee shop, suggesting that loyalty members are generally more supportive, although being a loyalty member does not guarantee a recommendation.

# **Insightful/Advanced Questions**

### **1. What factors (age, loyalty status, product_preference) are associated with higher satisfaction scores?**

```{r}
data_model <- data2 %>%
  select(Age, Loyalty_Member, Favorite_Product, Satisfaction_Score) %>%
  mutate(
    Loyalty_Member = factor(Loyalty_Member),
    Favorite_Product = factor(Favorite_Product)
  )

model <- lm(Satisfaction_Score ~ Age + Loyalty_Member + Favorite_Product, data = data_model)
summary(model)

```

```{r}
library(broom)

tidy(model) %>% 
  ggplot(aes(x = term, y = estimate, fill = p.value < 0.05)) +
  geom_col() +
  coord_flip() +
  labs(title = "Impact of Factors on Satisfaction Score",
       y = "Effect Size", x = "")
```

### **2. Is there a difference in satisfaction scores between customers who prefer coffee and those who prefer pastries?**

```{r}
coffee_pastries <- data2 %>% 
  filter(Favorite_Product %in% c('Coffee', 'Pastry')) %>% 
  group_by(Favorite_Product) %>% 
  summarise(Average_Satisfaction = mean(Satisfaction_Score)) %>% 
  ggplot(aes(x = Favorite_Product, y = Average_Satisfaction, fill = Favorite_Product)) +
  geom_col() +
  theme_minimal()

print(coffee_pastries)
```

### **3. Can we predict whether a customer would recommend the coffee shop based on other variables?**

```{r}
# Prepare data
data_model <- data2 %>%
  select(Would_Recommend, Gender, Satisfaction_Score, Age, Loyalty_Member, Visit_Frequency) %>%
  mutate(
    recommend = ifelse(Would_Recommend == "Yes", 1, 0),  
    Loyalty_Member = as.factor(Loyalty_Member),
    Gender = as.factor(Gender)
  ) %>%
  drop_na()
```

```{r}
# Fit logistic regression
log_model <- glm(recommend ~ Satisfaction_Score + Age + Loyalty_Member + Visit_Frequency + Gender, data = data_model, family = "binomial")

summary(log_model)
```

Yes, **we can partially predict** whether a customer would recommend the coffee shop:

-   **Satisfaction Score** is the strongest positive predictor (p = 0.043)

-   **Loyalty status** unexpectedly shows a negative effect (p = 0.025)

Other factors (age, gender, visit frequency) do not show significant effects in this model.
